---
title: '**Washtenaw County 65+ Demographic Report**'
author: "Washtenaw County Office of Community & Economic Development"
date: "`r format(Sys.Date(), format='%B %d, %Y')`"
output:
  bookdown::gitbook:
    config:
      download:
      - OCED-SNP_Report.pdf
    css: styles.css
    highlight: haddock
    number_sections: no
  pdf_document:
    fig_height: 3
    highlight: haddock
    latex_engine: xelatex
    number_sections: no
    toc: yes
  html_document:
    theme: cosmo
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      Message = FALSE, 
                      fig.margin = TRUE)

library(dplyr) # data management

library(tidyr) # tidying data

library(ggplot2) # beautiful grap?hs

library(ggthemes) # beautiful themes

library(plotly) # beautiful INTERACTIVE graphs

library(DT) # beautiful data tables

library(RColorBrewer)

library(ggbeeswarm)

library(png)

library(grid)


```

# Washtenaw County Meals on Wheels & Senior Cafe Program

The [Washtenaw County Meals on Wheels & Senior Cafe Program](https://www.washtenaw.org/snp) [(www.washtenaw.org/snp)](https://www.washtenaw.org/snp) is coordinated by the [Washtenaw County Office of Community and Economic Development](https://www.washtenaw.org/569/Community-Economic-Development) (OCED). OCED is committed to stepping out of traditional government roles to drive long-term system changes that increase equity and opportunity.

Please feel free to contact [Marti Lachapell](mailto:lachapellm@washtenaw.org), the Washtenaw County Meals on Wheels & Senior Café Program Specialist, if you have any questions.

<br><br>

Below is a diagram that indicates the role of OCED:

```{r}
library(DiagrammeR) # library for making diagrams

grViz("

digraph {

graph [rankdir = LR]

node [shape = oval, 
      style = filled, 
      fontname = Helvetica, 
      fontsize = 10, 
      fontcolor = white]

node [color = Goldenrod, fillcolor = Goldenrod]

OCED

node [color = SlateGray, fillcolor = SlateGray]
Federal State SeniorSites Seniors

edge [color = grey, arrowhead = none]
Federal -> OCED

State  -> OCED

OCED -> SeniorSites

SeniorSites -> Seniors

} ")
```

# Washtenaw County 65+ Seniors Report {.tabset}


*Note that Census Tracts which have no population for the race/ethnicities are eliminated.

*Data in this report are retrieved from the 2013-2017 American Community Survey 5-Year Estimates.*

\pagebreak


```{r}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

load ("~/Box Sync/shihyac/RData/WashtenawDemo.RData")

```


## 1. Comparison Graphs {.tabset}

### 1.1 Summary Table
```{r}

img8 <- readPNG ("~/Box Sync/shihyac/Tableau/Summary.png")

grid.raster(img8)
```

### 1.2 By Race/Ethnicity

For each race, every dot is a census tract.

```{r}

p1 <- Washtenaw_all %>%
  filter(Race != "Other" & Estimate != "0") %>%
  ggplot(aes(x = Race,
           y = Estimate, 
           color = Race, 
           fill = Race)) +
  labs(y = "Population Estimate")

p2 <- p1 + theme(axis.text.x = element_text(size = rel(1)), legend.position = "none") + 
  geom_boxplot(fill = "white") + labs(title = "", subtitle = "") + 
  xlab(" ") +
  scale_fill_tableau() +
  scale_color_tableau() + 
  theme_minimal()

ggplotly(p2) %>%
  layout(xaxis = list(title = ""),
         yaxis = list(title = "Population Estimate"), showlegend = FALSE) 

```

### 1.3 By Race/Ethnicity and Age

For each race, every dot is a census tract.

```{r}

p3 <- p1 +  theme(axis.text.x = element_text(size = rel(.8)), legend.position = "none") + 
  facet_wrap(~Age) + 
  geom_beeswarm() + 
  labs(title = "", subtitle = "") + xlab(" ") +
    scale_fill_tableau() +
  scale_color_tableau() + 
  theme_minimal()

ggplotly(p3, tooltip = "Estimate") %>%  
  layout(showlegend = FALSE)

```

\pagebreak

### 1.4 By Race

```{r}
pal <- c("blue", "orange", "red", "light blue", "dark green")

p_race2 <- Washtenaw_all %>% 
  group_by(Race) %>%
  filter(Race != "Other" & Race != "Pacific Islander" & !is.na(Estimate)) %>%
  summarise(Total_Population = sum(Estimate)) %>%
  plot_ly(type = 'scatter', mode = 'markers', 
          color = ~Race, colors = pal) %>%
  add_trace(x = ~Race, 
            y = ~Total_Population,
            marker = (list(
              size = 20))) %>%
  layout(title = "Comparison of Total Population by Race",
         xaxis = list(title = ""),
         yaxis = list(title = ""), showlegend = FALSE)


p_race2
```

### 1.5 By Gender

```{r out.width = "70%"}
img9 <- readPNG ("~/Box Sync/shihyac/Tableau/Gender.png")

grid.raster(img9)

```

## 2. Individual Race/Ethnicity Graphs {.tabset}

### 2.1 Asian
Each frequency indicates the amount of Census Tracts in Washtenaw County that have the according Asian population estimate in certain age group.

```{r}
p_asian <- Washtenaw_all %>%
  filter(Race == "Asian" & Estimate != "0") %>%
  ggplot(aes(x=Estimate,
             color = Gender, fill = Gender)) + 
  geom_histogram(alpha = .7) +
  facet_wrap(~Age) +
  labs(x = "Population Estimate", 
       y = "Frequency") +
  scale_fill_tableau() +
  scale_color_tableau()
  

ggplotly(p_asian, tooltip = c("Estimate", "count"))

```



### 2.2 Black
Each frequency indicates the amount of Census Tracts in Washtenaw County that have the according Black population estimate in certain age group.

```{r}

p_black <- Washtenaw_all %>%
  filter(Race == "Black" & Estimate != "0") %>%
  ggplot(aes(x=Estimate,
             color = Gender, fill = Gender)) + 
  geom_histogram(alpha = .7) +
  facet_wrap(~Age) + 
  labs(x = "Population Estimate", 
       y = "Frequency") +
  scale_fill_tableau() +
  scale_color_tableau()

ggplotly(p_black, tooltip = c("Estimate", "count"))

```

\pagebreak

### 2.3 Latinx
Each frequency indicates the amount of Census Tracts in Washtenaw County that have the according Latinx population estimate in certain age group.

```{r}

p_latinx <- Washtenaw_all %>%
  filter(Race == "Latinx" & Estimate != "0") %>%
  ggplot(aes(x=Estimate,
             color = Gender, fill = Gender)) + 
  geom_histogram(alpha = .7) +
  facet_wrap(~Age) + 
  labs(x = "Population Estimate", 
       y = "Frequency") +
    scale_fill_tableau() +
  scale_color_tableau()

ggplotly(p_latinx, tooltip = c("Estimate", "count"))

```

### 2.4 Pacific Islander
There is no 65+ seniors in Washtenaw County identified themselves as Pacific Islander in the *2013-2017 American Community Survey 5-Year Estimates*.

<br><br>
<br><br>
<br><br>
<br><br>


### 2.5 Native
Each frequency indicates the amount of Census Tracts in Washtenaw County that have the according Native population estimate in certain age group.

```{r}

p_native <- Washtenaw_all %>%
  filter(Race == "Native" & Estimate != "0") %>%
  ggplot(aes(x=Estimate,
             color = Gender, fill = Gender)) + 
  geom_histogram(alpha = .7) +
  facet_wrap(~Age) + 
  labs(x = "Population Estimate", 
       y = "Frequency") +
    scale_fill_tableau() +
  scale_color_tableau()

ggplotly(p_native, tooltip = c("Estimate", "count"))

```

\pagebreak

### 2.6 White
Each frequency indicates the amount of Census Tracts in Washtenaw County that have the according White population estimate in certain age group.

```{r}

p_white <- Washtenaw_all %>%
  filter(Race == "White" & Estimate != "0") %>%
  ggplot(aes(x=Estimate,
             color = Gender, fill = Gender)) + 
  geom_histogram(alpha = .7) +
  facet_wrap(~Age) + 
  labs(x = "Population Estimate", 
       y = "Frequency") +
    scale_fill_tableau() +
  scale_color_tableau()

ggplotly(p_white, tooltip = c("Estimate", "count"))

```


### 2.7 Other
Each frequency indicates the amount of Census Tracts in Washtenaw County that have the according other race/ethnicity population estimate in certain age group.

```{r}

p_other <- Washtenaw_all %>%
  filter(Race == "Other" & Estimate != "0") %>%
  ggplot(aes(x=Estimate,
             color = Gender, fill = Gender)) + 
  geom_histogram(alpha = .7) +
  facet_wrap(~Age) + 
  labs(x = "Population Estimate", 
       y = "Frequency") +
    scale_fill_tableau() +
  scale_color_tableau()

ggplotly(p_other, tooltip = c("Estimate", "count"))

```

\pagebreak


## 3. Demographics Maps {.tabset}

### 3.1 Asian
```{r}

img1 <- readPNG ("~/Box Sync/shihyac/QGIS/Exports/Individual Maps_Asian.png")

grid.raster(img1)
```

### 3.2 Black
```{r}

img2 <- readPNG ("~/Box Sync/shihyac/QGIS/Exports/Individual Maps_Black.png")

grid.raster(img2)
```

### 3.3 Latinx
```{r}

img3 <- readPNG ("~/Box Sync/shihyac/QGIS/Exports/Individual Maps_Latinx.png")

grid.raster(img3)
```


### 3.4 Native
```{r}

img4 <- readPNG ("~/Box Sync/shihyac/QGIS/Exports/Individual Maps_Native.png")

grid.raster(img4)
```

### 3.5 White
```{r}

img5 <- readPNG ("~/Box Sync/shihyac/QGIS/Exports/Individual Maps_White.png")

grid.raster(img5)
```

### 3.6 Other Race
```{r}

img6 <- readPNG ("~/Box Sync/shihyac/QGIS/Exports/Individual Maps_Other.png")

grid.raster(img6)
```

### 3.7 All Races
```{r}

img7 <- readPNG ("~/Box Sync/shihyac/QGIS/Exports/Washtenaw County 65+ Population Demographics.png")

grid.raster(img7)
```