---
title: '**Washtenaw County 65+ Demographic Report**'
author: "Washtenaw County Office of Community & Economic Development"
date: "`r format(Sys.Date(), format='%B %d, %Y')`"
output:
  bookdown::gitbook:
    config:
      download:
      - Washtenaw-County-Senior-Demographic-Report.pdf
      - Washtenaw-County-Senior-Demographic-Report.epub
    css: styles.css
    highlight: haddock
    number_sections: no
  html_document:
    theme: cosmo
    toc: no
  pdf_document:
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      Message = FALSE, 
                      fig.margin = TRUE)

library(ggplot2)

library(dplyr)

library(ggthemes)

library(ggbeeswarm)


```

# Washtenaw County Meals on Wheels & Senior Cafe Program

The [Washtenaw County Meals on Wheels & Senior Cafe Program](https://www.washtenaw.org/snp) [(www.washtenaw.org/snp)](https://www.washtenaw.org/snp) is coordinated by the [Washtenaw County Office of Community and Economic Development](https://www.washtenaw.org/569/Community-Economic-Development) (OCED). OCED is committed to stepping out of traditional government roles to drive long-term system changes that increase equity and opportunity.

Please feel free to contact [Marti Lachapell](mailto:lachapellm@washtenaw.org), the Washtenaw County Meals on Wheels & Senior Café Program Specialist, if you have any questions.

# Washtenaw County 65+ Seniors Report {.tabset}

Please click the tabs below to see graphs of Washtenaw County 65+ seniors demographics by race/ethnicity and age groups. 

*Note that Census Tracts which have no population for the race/ethnicities are eliminated.

*Data in this report are retrieved from the 2013-2017 American Community Survey 5-Year Estimates.*


```{r}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

load ("~/Box Sync/shihyac/RData/WashtenawDemo.RData")

```

## 1. Individual Race/Ethnicity Graphs {.tabset}

### 1.1 Asian
Each frequency indicates the amount of Census Tracts in Washtenaw County that have the according Asian population estimate in certain age group.

```{r}
Washtenaw_all %>%
  filter(Race == "Asian" & Estimate != "0") %>%
  ggplot(aes(x=Estimate,
             color = Gender, fill = Gender)) + 
  geom_histogram(alpha = .7) +
  facet_wrap(~Age) + 
  labs(x = "Population Estimate", 
       y = "Frequency") +
  scale_fill_tableau() +
  scale_color_tableau()

```



### 1.2 Black
Each frequency indicates the amount of Census Tracts in Washtenaw County that have the according Black population estimate in certain age group.

```{r}

Washtenaw_all %>%
  filter(Race == "Black" & Estimate != "0") %>%
  ggplot(aes(x=Estimate,
             color = Gender, fill = Gender)) + 
  geom_histogram(alpha = .7) +
  facet_wrap(~Age) + 
  labs(x = "Population Estimate", 
       y = "Frequency") +
  scale_fill_tableau() +
  scale_color_tableau()

```

### 1.3 Latinx
Each frequency indicates the amount of Census Tracts in Washtenaw County that have the according Latinx population estimate in certain age group.

```{r}

Washtenaw_all %>%
  filter(Race == "Latinx" & Estimate != "0") %>%
  ggplot(aes(x=Estimate,
             color = Gender, fill = Gender)) + 
  geom_histogram(alpha = .7) +
  facet_wrap(~Age) + 
  labs(x = "Population Estimate", 
       y = "Frequency") +
    scale_fill_tableau() +
  scale_color_tableau()

```

### 1.4 Pacific Islander
There is no 65+ seniors in Washtenaw County identified themselves as Pacific Islander in the *2013-2017 American Community Survey 5-Year Estimates*.

<br><br>
<br><br>
<br><br>
<br><br>
<br><br>
<br><br>
<br><br>
<br><br>
<br><br>
<br><br>
<br><br>

### 1.4 Native
Each frequency indicates the amount of Census Tracts in Washtenaw County that have the according Native population estimate in certain age group.

```{r}

Washtenaw_all %>%
  filter(Race == "Native" & Estimate != "0") %>%
  ggplot(aes(x=Estimate,
             color = Gender, fill = Gender)) + 
  geom_histogram(alpha = .7) +
  facet_wrap(~Age) + 
  labs(x = "Population Estimate", 
       y = "Frequency") +
    scale_fill_tableau() +
  scale_color_tableau()

```

### 1.5 White
Each frequency indicates the amount of Census Tracts in Washtenaw County that have the according White population estimate in certain age group.

```{r}

Washtenaw_all %>%
  filter(Race == "White" & Estimate != "0") %>%
  ggplot(aes(x=Estimate,
             color = Gender, fill = Gender)) + 
  geom_histogram(alpha = .7) +
  facet_wrap(~Age) + 
  labs(x = "Population Estimate", 
       y = "Frequency") +
    scale_fill_tableau() +
  scale_color_tableau()

```

### 1.6 Other
Each frequency indicates the amount of Census Tracts in Washtenaw County that have the according other race/ethnicity population estimate in certain age group.

```{r}

Washtenaw_all %>%
  filter(Race == "Other" & Estimate != "0") %>%
  ggplot(aes(x=Estimate,
             color = Gender, fill = Gender)) + 
  geom_histogram(alpha = .7) +
  facet_wrap(~Age) + 
  labs(x = "Population Estimate", 
       y = "Frequency") +
    scale_fill_tableau() +
  scale_color_tableau()

```

## 2. Comparison Graphs {.tabset}

### 2.1 By Race/Ethnicity
```{r}

p1 <- Washtenaw_all %>%
  filter(Race != "Other" & Estimate != "0") %>%
  ggplot(aes(x = Race,
           y = Estimate, 
           color = Race, 
           fill = Race)) +
  labs(y = "Population Estimate")

p1 + theme(axis.text.x = element_text(size = rel(1))) + 
  geom_boxplot(fill = "white") + labs(title = "Comparison by Race/Ethnicity", subtitle = "For each race every dot is a census tract") + xlab(" ") +
    scale_fill_tableau() +
  scale_color_tableau()

```

### 2.2 By Race/Ethnicity and Age

```{r}

p1 +  theme(axis.text.x = element_text(size = rel(.8))) + 
  facet_wrap(~Age) + geom_beeswarm() + labs(title = "Comparison by Race/Ethnicity and Age", subtitle = "For each group every dot is a census tract") + xlab(" ") +
    scale_fill_tableau() +
  scale_color_tableau()

```

### 2.3 Total Population

```{r}
facet_horiz_dotchart <- Washtenaw_all %>%
  filter(Race != "Other" & Estimate != "0") %>%
  ggplot(aes(x = factor(Race),y = Estimate, color = Race)) + 
          stat_summary(fun.y = sum, 
          geom = "point", 
          size = 7) +
  guides(fill=FALSE) +
  xlab(" ") + 
  ylab("Total Population Estimate") + 
  labs(title = "Comparison of Total Population by Race/Ethnicity") +
  theme(axis.title.y = element_blank(),
        axis.ticks = element_blank()) +
    scale_fill_tableau() +
  scale_color_tableau()

facet_horiz_dotchart
```

